---
title: "mOMOP to OMOP Vocabularies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mOMOP to OMOP Vocabularies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
```

```{r setup}
library(amphora)
library(mOMOP)
library(tidyverse)
```

In lieu of using Protege, mCode value set classification developed in the mOMOP
package is converted to the OMOP Vocabulary structure, specifically the class 
hierarchy seen in the Concept Relationship table. This is achieved by creating:  
1. Forward "Subsumes" relationships between the `class` and `concept`.  
1. Inverse "Is a" relationships between the `concept` and class.  
1. Concepts for each `class` in a Concept table. 

```{r}
MCODE_CLASS_HIERARCHY
```

The `class` field is converted into a human readable format. 

```{r}
MCODE_CLASS_HIERARCHY2 <-
  MCODE_CLASS_HIERARCHY %>%
  mutate_at(vars(class), stringr::str_replace_all, "_", " ") %>%
  mutate_at(vars(class), stringr::str_to_title)
unique(MCODE_CLASS_HIERARCHY2$class)
```

```{r}
CONCEPT_A <-
  MCODE_CLASS_HIERARCHY2 %>%
  select(class) %>%
  distinct() %>%
  mutate(concept_id = uuid::UUIDgenerate(n = length(unique(MCODE_CLASS_HIERARCHY2$class))),
         concept_code = length(unique(MCODE_CLASS_HIERARCHY2$class))) %>%
  transmute(
    concept_id,
    concept_name = class,
    domain_id = "Cancer Modifier",
    vocabulary_id = "mOMOP",
    concept_class_id = "Class",
    standard_concept = "C",
    concept_code,
    valid_start_date = Sys.Date(),
    valid_end_date = as.Date(NA_character_),
    invalid_reason = NA_character_)
rMarkedDown::print_dt(CONCEPT_A)
```

```{r}
CONCEPT_RELATIONSHIP <-
CONCEPT_A %>%
  transmute(concept_id_1 = concept_id,
            class = concept_name) %>%
  inner_join(MCODE_CLASS_HIERARCHY2,
            by = "class") %>%
  chariot::unmerge_strip(strip_col = concept,
                         remove = FALSE) %>%
  mutate(concept_id_2 = coalesce(concept_id, concept)) %>%
  transmute(concept_id_1,
             relationship_id = "Is a",
             concept_id_2,
             valid_start_date = Sys.Date(),
             valid_end_date = as.Date(NA_character_),
             invalid_reason = NA_character_)
```


```{r,message='hide'}
CONCEPT_B <-
chariot::join_on_concept_id(
  data = CONCEPT_RELATIONSHIP %>%
            mutate_at(vars(concept_id_2), as.integer),
  column = "concept_id_2"
) %>%
  select(concept_id:last_col())
```

Both subsets of the Concept data, including `concept_id` is converted to character because the `uuid` identifiers created are not integers. 

```{r}
CONCEPT <-
  bind_rows(CONCEPT_A %>%
              mutate_all(as.character),
            CONCEPT_B %>%
              mutate_all(as.character))
```

The dataset is written to data-raw/ if it does not already exist. 

```{r}
MOMOP_TABLES <-
  list(CONCEPT = CONCEPT,
       CONCEPT_RELATIONSHIP = CONCEPT_RELATIONSHIP) 

file <- file.path("data-raw", "MOMOP_TABLES.xlsx")
if (!file.exists(file)) {
  broca::write_full_excel(x = MOMOP_TABLES,
                          file = file)
}
```

