---
title: "Visualize False Positive Antineoplastic Drug Classes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualize False Positive Antineoplastic Drug Classes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "##",
                      tidy = TRUE, #`styler` to use styler:style_text() to reformat code
                      tidy.opts = list(blank = FALSE, width.cutoff = 60),
                      echo = TRUE,
                      eval = TRUE,
                      cache = TRUE,
                      # cache.path = "reports/cache/",
                      child = NULL, #file/s to knit and then include,
                      collapse = FALSE, #collapse all output into a single block,
                      error = TRUE, #display error messages in doc. FALSE stops render when error is thrown
                      fig.align = "center", #left, right, center, or default
                      fig.width = 7, #inches
                      fig.height = 7, #inches
                      include = TRUE, #include chunk?
                      message = FALSE, #display code messages?
                      tidy = TRUE, #tidy code 
                      warning = FALSE, #include warnings?
                      results = "markup"
                        # "asis": passthrough results
                        # "hide": do not display results 
                        # "hold": put all results below all code
                      )
```
  
```{r libraries}
library(tidyverse)
library(pg13)
library(chariot)
library(amphora)
```

  
```{r,results='hide'}
conn <- connectAthena()
```

There are drugs that are used to treat both cancerous and non-cancerous conditions. 
Therefore, it cannot be reasonably assumed that the patient is a cancer patient
if these `false positive drugs` are present in the drug exposure record. Below is 
a curated list of the RxNorm Ingredients that qualify as `false positive drugs` 
according to feedback from a panel of Oncologists at OHDSI. 

```{r false_positive_concepts0,echo=FALSE,results='hide'}
false_positive_ingrs <-
tibble::tribble(~rxnorm_ingredient_name,
        "filgrastim",
        "benralizumab",
        "cyclosporine",
        "dihematoporphyrin ether",
        "eculizumab",
        "ethoglucid",
        "fosfestrol",
        "fostamatinib",
        "herpesvirus 1, human",
        "mesna",
        "pravastatin",
        "ravulizumab",
        "ruxolitinib",
        "tranexamic acid",
        "valproate",
        "levamisole",
        "methyl 5-aminolevulinate",
        "phenytoin",
        "zidovudine",
        "mepolizumab",
        "mycophenolate mofetil",
        "ketoconazole",
        "pirfenidone",
        "sargramostim",
        "octreotide",
        "quinine",
        "tacrolimus",
        "megestrol",
        "miltefosine",
        "medroxyprogesterone",
        "methylprednisolone",
        "omeprazole",
        "tocilizumab",
        "leucovorin",
        "levoleucovorin",
        "hydrocortisone",
        "cortisone",
        "danazol",
        "doxycycline",
        "estradiol",
        "dexamethasone",
        "denosumab",
        "amoxicillin",
        "aspirin",
        "azathioprine",
        "sirolimus",
        "isotretinoin",
        "diethylstilbestrol",
        "dutasteride",
        "prednisolone",
        "prednisone",
)

false_positive_concepts1 <-
join_on_concept_synonym_name(
  data = false_positive_ingrs,
  column = "rxnorm_ingredient_name",
  conn = conn
) %>%
  rename(match_concept_id = concept_id)

false_positive_concepts <-
  join_on_concept_id(
    data = false_positive_concepts1,
    column = "match_concept_id",
    where_in_concept_field = "vocabulary_id",
    where_in_concept_field_value = c("RxNorm", "RxNorm Extension"),
    where_is_null_concept_field = "invalid_reason",
    conn = conn
  ) %>%
  select(concept_id:invalid_reason)
```

```{r false_positive_concepts}
false_positive_concepts
```

Reverse engineering the false positives will allows for identification of the 
OMOP Concept Classes in ATC and/or HemOnc that contain the `false positive drugs` 
that can then be excluded from analyses. This entails finding the immediate parent 
class of each of the RxNorm Ingredient identified as a `false positive drug` using 
relationships with a maximum level of separation of 1 in the Concept Ancestor table.   

```{r,results='hide'}
false_positive_ancestors <-
join_for_ancestors(
  data = false_positive_concepts,
  descendant_id_column = "concept_id",
  where_in_concept_ancestor_field = "max_levels_of_separation",
  where_in_concept_ancestor_field_value = 1,
  conn = conn
  )
```

Additional filters are applied to the parents of the `false positive drugs`:  
1. OMOP Classes  
1. ATC and HemOnc Vocabularies because they are the target classifications  
1. Concept Classes not equal to ATC 5th because it has a lateral relationship to 
RxNorm Ingredient Concepts

```{r false_positive_parents}
false_positive_parents <-
  false_positive_ancestors %>%
  rename_all(str_replace_all, "ancestor_", "parent_") %>%
  filter(
    parent_vocabulary_id %in% c("ATC", "HemOnc"),
    parent_standard_concept %in% c("C")
    )  %>%
  filter(!(parent_concept_class_id %in% c("ATC 5th")))
false_positive_parents
```

The classifications for ATC and HemOnc are separated from each other to analyze 
separately, starting with ATC to return ATC Classes that contain 
`false positive drugs`. A `color` column is added to direct `fill` in the plot.

```{r atc}
atc <-
false_positive_parents %>%
  rubix::split_by(col = parent_vocabulary_id) %>%
  pluck("ATC") %>%
  unite(col = false_positive_atc_class,
        parent_concept_id,
        parent_concept_name,
        sep = " ") %>%
  transmute(false_positive_atc_class,
            color = "red")
atc
```

To visualize these ATC Classes, the class hierarchy of the 
`ANTINEOPLASTIC AND IMMUNOMODULATING AGENTS` topmost ATC class is derived. 

```{r,results='hide'}
atc_descendants <-
  join_for_descendants(data = tibble::tibble(top_atc_class = 21601386)) %>%
  filter(descendant_vocabulary_id %in% "ATC") %>%
  filter(!(descendant_concept_class_id %in% "ATC 5th")) %>%
  arrange(descendant_concept_class_id) %>%
  unite(col = descendant,
        descendant_concept_id,
        descendant_concept_name,
        sep = " ",
        remove = FALSE) %>%
  pivot_wider(id_cols = top_atc_class,
              names_from = descendant_concept_class_id,
              values_from = descendant,
              values_fn = list(descendant = ~ paste(unique(.), collapse = "|"))) %>%
  rubix::recode_value(col = top_atc_class,
                      "21601386 ANTINEOPLASTIC AND IMMUNOMODULATING AGENTS" = "21601386")
```

```{r atc_descendants}
atc_descendants
```

The dataframe of the ATC class hierarchy for 
`ANTINEOPLASTIC AND IMMUNOMODULATING AGENTS` is depivoted into a parent-child 
data tree structure. A `color` field is added to plot all classes in blue unless 
it is a class containing a `false positive drug`, in which case it will be plotted 
in red. 

```{r}
parent_child_fields <- 
  list(
    c("top_atc_class", "ATC 1st"),
    c("ATC 1st", "ATC 2nd"),
    c("ATC 2nd", "ATC 3rd"),
    c("ATC 3rd", "ATC 4th")
  )

output <- list()
for (i in seq_along(parent_child_fields)) {
  x <-
  atc_descendants %>%
    select(all_of(parent_child_fields[[i]])) 
  colnames(x) <- c("parent", "child")
  
  output[[i]] <- x
  
}
atc_data_tree <-
  bind_rows(output) %>%
  separate_rows(parent, sep = "[|]{1}") %>%
   separate_rows(child, sep = "[|]{1}") %>%
  mutate(color = "blue") %>%
  distinct()
atc_data_tree
```

The ATC classes containing the `false positive drugs` is joined to the data tree 
to denote them in red.

```{r}
atc_data_tree2 <-
atc_data_tree %>%
  left_join(atc, 
            by = c("child" = "false_positive_atc_class"),
            suffix = c(".dt", ".fp")) %>%
  transmute(parent,
            child,
            color = coalesce(color.fp, color.dt))
atc_data_tree2
```

The root node is converted to `NA_character_` as a requirement for the 
`collapsibleTreeNetwork` function in the `collapsibleTree` package.  

```{r}
atc_data_tree2$parent[1]<- NA_character_
atc_data_tree2
```

### ATC Plot  

```{r,cache=FALSE}
collapsibleTreeNetwork(df = atc_data_tree2,
                       fill = "color")
```



```{r,results='hide'}
dcAthena(conn = conn)
```

