---
title: 'Methods: Reverse Engineering False Positives'
author: "Meera Y. Patel, M.D."
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    highlight: kate  
    code_folding: show #or hide
    df_print: paged
---
```{r, include = FALSE}
knitr::opts_chunk$set(comment = "##",
                      tidy = TRUE, #`styler` to use styler:style_text() to reformat code
                      tidy.opts = list(blank = FALSE, width.cutoff = 60),
                      echo = TRUE,
                      eval = TRUE,
                      cache = TRUE,
                      # cache.path = "reports/cache/",
                      child = NULL, #file/s to knit and then include,
                      collapse = FALSE, #collapse all output into a single block,
                      error = TRUE, #display error messages in doc. FALSE stops render when error is thrown
                      include = TRUE, #include chunk?
                      message = FALSE, #display code messages?
                      tidy = TRUE, #tidy code 
                      warning = FALSE, #include warnings?
                      results = "markup"
                        # "asis": passthrough results
                        # "hide": do not display results 
                        # "hold": put all results below all code
                      )
```


```{r}
library(tidyverse)
library(pg13)
library(ggplot2)
library(amphora)
conn <- pg13::local_connect()
```

Reverse engineering the false positives will for identification of the 
OMOP Concept Classes in ATC and/or HemOnc that contain the false positive drugs 
that can then be excluded from analyses. 298 RxNorm Ingredients falling under 
the ATC Antineoplastics Class were reviewed and drugs that are less likely to be 
used in an oncology context and/or is more likely to be used for treatment of a 
disease unrelated to cancer were identified by a panel of oncologists, the 
results of which are included in the `antineoplastics_annotation` dataset in 
this package. 

<br>

```{r}
rMarkedDown::print_dt(antineoplastics_annotation)
```

<br>

The annotated dataframe is filtered for the `false positive` RxNorm Ingredients.  

```{r}
false_positives <-
  antineoplastics_annotation %>%
  filter(false_positive == TRUE)
false_positives
```

The `false positives` are first rejoined with the Concept table to retrieve all 
the concept attributes. 

```{r}
false_positives2 <-
  chariot::join_on_concept_id(
    data = false_positives %>%
              select(false_positive_id = concept_id)
  ) %>%
  select(-false_positive_id)
false_positives2
```

The classification of the false positives is derived from the Concept Ancestor 
Table, where any ancestors to the false positives are first identified and 
selected for if they are a OMOP Drug Class.  

```{r}
false_positive_ancestors <-
  false_positives2 %>%
  left_join(
    chariot::join_for_ancestors(data = false_positives2,
                                descendant_id_column = "concept_id") %>%
      # Filter for classes 
      filter(ancestor_standard_concept == "C") %>%
      # Filter for valid concepts
      filter(is.na(ancestor_invalid_reason))  %>%
      filter(ancestor_vocabulary_id %in% c("HemOnc", "ATC")),
    by = c(
        "concept_id", 
           "concept_name", 
           "domain_id", 
           "vocabulary_id", 
           "concept_class_id", 
           "standard_concept", 
           "concept_code", 
           "valid_start_date", 
           "valid_end_date", 
           "invalid_reason"
        )
    ) %>%
  distinct()
false_positive_ancestors
```

The following special relationships require additional modification before
moving forward:

@  There may be cases where the minimum level of separation is not equal
    to the max levels of separation. Since this particular use case indicates an
    offset of 1 level, the maximum level of separation is used to determine the
    hierarchy.

```{r}
false_positive_ancestors %>%
        filter(min_levels_of_separation != max_levels_of_separation) %>%
        select(ancestor_vocabulary_id,
               ancestor_concept_class_id,
               min_levels_of_separation, 
               max_levels_of_separation) %>%
        distinct()
```

@ There are cases of O levels of separation between an RxNorm Ingredient and 
either a ATC or a HemOnc Class. These cases will be treated as 1 level of 
separation.   

```{r}
false_positive_ancestors %>%
        filter(min_levels_of_separation == 0|max_levels_of_separation == 0) %>%
        select(ancestor_vocabulary_id,
               ancestor_concept_class_id,
               min_levels_of_separation, 
               max_levels_of_separation) %>%
        distinct()
```

The resolve the scenarios outlined:  
* All ATC 5th Concept Classes are excluded because they are redundant to  
the RxNorm Ingredient and therefore do not add clinical value  
* `min_levels_of_separation` as the source of hierarchy 

```{r}
false_positive_ancestors2 <-
  false_positive_ancestors %>%
  filter(!(ancestor_concept_class_id %in% c("ATC 5th"))) %>%
  mutate(level_of_separation = max_levels_of_separation) %>%
  select(-min_levels_of_separation, -max_levels_of_separation) %>%
  chariot::merge_strip(into = "ancestor", ancestor_vocabulary_id, prefix = "ancestor_") %>%
  chariot::merge_strip(into = "concept") %>%
  select(-concept_id, -ancestor_id)
```

# HemOnc 

```{r}
false_positive_ancestors2_ho <-
  false_positive_ancestors2 %>%
  filter(ancestor_vocabulary_id %in% "HemOnc") %>%
  select(-ancestor_vocabulary_id)
```


```{r}
false_positive_ancestors2_ho2 <-
false_positive_ancestors2_ho %>%
        pivot_wider(id_cols = concept,
                    names_from = level_of_separation,
                    values_from = ancestor,
                    values_fn = function(x) paste(unique(x), collapse = "|")) %>%
        select(concept, all_of(as.character(0:4))) %>%
        separate_rows(`1`, sep = "[|]{1}") %>%
        separate_rows(`2`, sep = "[|]{1}") %>%
        separate_rows(`3`, sep = "[|]{1}") %>%
        separate_rows(`4`, sep = "[|]{1}")
```

Since there are some NA values at levels, they are replaced with a dummy value. 

```{r}
false_positive_ancestors2_ho2 %>%
  filter_all(any_vars(is.na(.)))
```


```{r}
parent_child_cols <-
  list(c(parent = "concept", child = "0"),
       c(parent = "0", child =  "1"),
       c(parent = "1", child =  "2"),
       c(parent = "2", child =  "3"),
       c(parent = "3", child =  "4"))

output <- list()
for (i in seq_along(parent_child_cols)) {
  parent_child_col <- parent_child_cols[[i]]
  child_col <- sym(parent_child_col["child"])
  parent_col <- sym(parent_child_col["parent"])
  col_to_coalesce <- sym(as.character(as.integer(parent_child_col["child"])+1))
  
  output[[i]] <-
    false_positive_ancestors2_ho2 %>%
    transmute(parent = {{ parent_col }},
              child = coalesce({{ child_col }}, {{ col_to_coalesce }})) 
  
  colnames(output[[i]]) <- c("parent", "child")
  
}
false_positive_ancestors2_ho3 <-
  bind_rows(output) %>%
  distinct()
false_positive_ancestors2_ho3
```

```{r}
  if (i == 1) {
    root <- 
      tibble(parent = NA_character_,
                child = "HemOnc")
    root2 <-
      output[[i]] %>%
      mutate(child = parent) %>%
      mutate(parent = "HemOnc") %>%
      select(parent, child)
  }
```





```{r}
false_positive_ancestors2_ho4 <-
  false_positive_ancestors2_ho3 %>%
  chariot::unmerge_strip(strip_col = child)
```


```{r}
ho_tree <- collapsibleTree::collapsibleTreeNetwork(
                                false_positive_ancestors2_ho3,
                                 zoomable = TRUE,
                                 collapsed = FALSE
                                 )
```

```{r}

htmlwidgets::saveWidget(widget = ho_tree,
                        file = file.path(getwd(), "vignettes/widgets/reverse_engineered_false_positives_ho.html"))
```

# ATC  

```{r}
false_positive_ancestors2_atc <-
  false_positive_ancestors2 %>%
  filter(ancestor_vocabulary_id %in% "ATC") %>%
  select(-ancestor_vocabulary_id)
```


Though ATC has 5 levels, only 4 are chosen since the topmost level (5 in this use case) is too general for us to use as a blanket exclusion criteria.  

```{r}
atc_collapsible_tree <-
false_positive_ancestors2_atc  %>%
        arrange(level_of_separation) %>%
        pivot_wider(id_cols = concept,
                    names_from = level_of_separation,
                    values_from = ancestor,
                    values_fn = function(x) paste(unique(x), collapse = "|")) %>%
        select(concept, `1`, `2`, `3`) %>%
        separate_rows(`1`, sep = "[|]{1}") %>%
        separate_rows(`2`, sep = "[|]{1}") %>%
        separate_rows(`3`, sep = "[|]{1}") %>%
        distinct()
```


```{r}
atc_tree <- collapsibleTree::collapsibleTree(df = atc_collapsible_tree, 
                                 hierarchy = c("1", "2", "3"),
                                 root = "ATC False Positives",
                                 linkLength = 350,
                                 width = 1250,
                                 height = 22000,
                                 zoomable = TRUE,
                                 collapsed = FALSE)
htmlwidgets::saveWidget(widget = atc_tree,
                        file = file.path(getwd(), "vignettes/widgets/reverse_engineered_false_positives_atc.html"))
```


```{r}
# pg13::drop_all_staging_tables(conn = conn,
#                               schema = "public")
pg13::dc(conn = conn)
```

